Structural Breaks in Inflation Dynamics within the European Monetary Union

### 2010-10-08 pkg/paper/Draft/draft-Z-W.tex
small touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@38 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@

Another possible explanation pattern could be the growth in the  money supply. A drawback is the current lack of data\footnote{hopefully to be solved}. But for some countries we can find some plausible findings. Regarding the case of Denmark, the M1 money supply started contracting in June 2000 (on a monthly base). In Estonia we find the highpoint in M3 expansion ending with March 1998, with much lower M3 growth rates afterwards. The same happended to Hungary in April and May 1998. In the Polnish case we find a severe contraction of the money supply in May 2001. Similar findings hold true for Germany as well. \\
In Spain we see a stronger increase in M1 in May and June 1996 (nothing for the other break, data scarceity). In Portugal we find a modest increase in year on year M1 money supply growth starting in July 1992. We find nothing peculiar for Slovakia and Hungary. In Estonia the M1 money supply starts to contract severely on a yearly base following March 1998. In Slovenia we find a contraction of the Money Supply in July 2003, with the whole year 2003 witnessing a sharp decrease in the year on year growth. Poland starts contracting the year on year M1 money supply after June 2000 with very low growth rates till January 2002. 
\\
[irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck]

@@

Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime. \\

With regard to the alleged benefit of monetary unions, the import of credibility to former high inflation countries (like Italy and Spain), did nothing to decrease their mean or volatility, rather to the contrary. Spain and Italy did have a -- rather -- low mean low volatility regime from 1996:6 to 2000:12 but then things changed towards higher volatility. It appears as if both countries tried to decrease inflation prior to the fixing of the exchange rates and let go afterwards.  \\

Following the reasoning of a study by \cite{duarte}, high volatility of inflation rates within regions of a union may be explained a countries decision to lower the inflation differential by means of fiscal policy. This explanation might fit to Spain and Italy, both were above the the average EU inflation. \\

>>>>

%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}

@@

Another possible explanation pattern could be the growth in the  money supply. A drawback is the current lack of data\footnote{hopefully to be solved}. But for some countries we can find some plausible findings. Regarding the case of Denmark, the M1 money supply started contracting in June 2000 (on a monthly base). In Estonia we find the highpoint in M3 expansion ending with March 1998, with much lower M3 growth rates afterwards. The same happended to Hungary in April and May 1998. In the Polnish case we find a severe contraction of the money supply in May 2001. Similar findings hold true for Germany as well. \\
In Spain we see a stronger increase in M1 in May and June 1996 (nothing for the other break, data scarceity). In Portugal we find a modest increase in year on year M1 money supply growth starting in July 1992. We find nothing peculiar for Slovakia and Hungary. In Estonia the M1 money supply starts to contract severely on a yearly base following March 1998. In Slovenia we find a contraction of the Money Supply in July 2003, with the whole year 2003 witnessing a sharp decrease in the year on year growth. Poland starts contracting the year on year M1 money supply after June 2000 with very low growth rates till January 2002.

\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}

@@

Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime.

With regard to the alleged benefit of monetary unions, the import of credibility to former high inflation countries (like Italy and Spain), did nothing to decrease their mean or volatility, rather to the contrary. Spain and Italy did have a -- rather -- low mean low volatility regime from 1996:6 to 2000:12 but then things changed towards higher volatility. It appears as if both countries tried to decrease inflation prior to the fixing of the exchange rates and let go afterwards.

Following the reasoning of a study by \cite{duarte}, high volatility of inflation rates within regions of a union may be explained a countries decision to lower the inflation differential by means of fiscal policy. This explanation might fit to Spain and Italy, both were above the the average EU inflation.
